---
title: "Iteration and listcols"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(rvest)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Lists

You can ut anything in a list.

```{r}
## concept: lists can store anything.
vec_numeric = 5:8
vec_logical = c(TRUE, TRUE, FALSE, TRUE, FALSE, FALSE)
mat = matrix(1:8, nrow = 2, ncol = 4)
summary = summary(rnorm(100))

##these are all different kinds of lists, different contents and forms
##can't put these in a dataframe but can put in a list

l = list(
  vec_numeric = 5:8,
  vec_logical = c(TRUE, TRUE, FALSE, TRUE, FALSE, FALSE),
  mat = matrix(1:8, nrow = 2, ncol = 4),
  summary = summary(rnorm(100))
)
```

```{r}
l
l$vec_numeric
l[[1]]
l[["vec_numeric"]]
l[["vec_numeric"]][1:3]
mean(l[["vec_numeric"]])
## you can do whatevr you normally do to individual elements of a list
```

## `for` loop

Create a new list.

```{r}
list_norm = 
  list(
    a = rnorm(20, mean = 3, sd = 1),
    b = rnorm(30, mean = 0, sd = 5),
    c = rnorm(40, mean = 10, sd = .2),
    d = rnorm(20, mean = -3, sd = 1)
  )
```

```{r}
list_norm[[1]]
list_norm
```

Pause and get my old function.

```{r}
mean_and_sd = function(x) {
  
  if (!is.numeric(x)) {
    stop("Input must be numeric")
  }
  
  if (length(x) < 3) {
    stop("Input must have at least three numbers")
  }

  mean_x = mean(x)
  sd_x = sd(x)
 
  tibble(
    mean = mean_x,
    sd = sd_x
  )
  
}
```

I can apply that function to each list element.

```{r}
list_norm[[1]]
list_norm[[2]]
list_norm[[3]]
list_norm[[4]]

mean_and_sd(list_norm[[1]])
mean_and_sd(list_norm[[2]])
mean_and_sd(list_norm[[3]])
mean_and_sd(list_norm[[4]])

## so applying a function to each of my list items - that's fine.  With 4 lists, that's ok, but if 40 lists, more tedious. so, let's use a for loop.
```

Let's use a for loop:

```{r}
output = vector("list", length = 4)

for (i in 1:4) {
  
  output[[i]] = mean_and_sd(list_norm[[i]])

}
```


## Let's try map!
Instead of saying input list and output list and then do the procedure, the idea behind map is ...


```{r}
output = map(list_norm, mean_and_sd)
## doing same thing and even saving the list now
```

what if you want a different function..?

```{r}
output = map(list_norm, median)
output = map(list_norm, IQR)
## you can map any function you want across this input list
## this is a function within a function, whatever you want
```

```{r}
output = map(list_norm, median)

output = map_dbl(list_norm, median)
## this creates a different kind of fancier list - it is giving e a vector
```

```{r}
output = map_df(list_norm, mean_and_sd)
## what this does is instead of having multiple lists (4 lists), it creates a dataframe and binds them 
output = map_df(list_norm, mean_and_sd, .id = "input")
## here, .id = "input" tells it to keep track of from which list the mean and sd are coming from
```

Up to this point, for loops, map statements, we got lists.
Key idea is keeping track of everything so in your environment, you've got a lot of stuff.
What you want to have is something that keeps tracks of inputs and outputs at the same time.

## List columns!

```{r}
listcol_df =
  tibble(
    name = c("a", "b", "c", "d"),
    samp = list_norm
  )
## this is a dataframe - printed tibble shows you the name and the lists and their elements
```

```{r}
listcol_df %>% pull(name)
listcol_df %>% pull(samp)
## so the listcol_df still has all the elements that should be there, it is just a more compact way of looking at things
## it's a dataframe, so you can still do whatever you want to a dataframe

listcol_df %>%
  filter(name == "a")


```

Let's try some operations.

```{r}
listcol_df$samp
## dollar sign is the same thing as pulling out that list - a little easier in this example

listcol_df$samp[[1]]

mean_and_sd(listcol_df$samp[[1]])
## so we are applying mean_and_sd function to the df and only the first element in that dataframe

mean_and_sd(listcol_df$samp[[2]])
mean_and_sd(listcol_df$samp[[3]])
mean_and_sd(listcol_df$samp[[4]])
```

Can I just map??

```{r}
map(listcol_df$samp, mean_and_sd)

```

So... can I add a list column??
YES
Mutate issue!
```{r}
listcol_df %>% 
  mutate(summary = map(samp, mean_and_sd))

listcol_df =
  listcol_df %>% 
  mutate(summary = map(samp, mean_and_sd))

listcol_df =
  listcol_df %>% 
  mutate(summary = map_df(samp, mean_and_sd))

listcol_df =
  listcol_df %>% 
  mutate(
    summary = map(samp, mean_and_sd),
    medians = map(samp, median)
  )

listcol_df =
  listcol_df %>% 
  mutate(
    summary = map(samp, mean_and_sd),
    medians = map_dbl(samp, median)
  )

## so the process here, in a few lines of code, is to say i hae a df, created a mean_and_sd summary, and median for these elements... you don't have a bunch of vectors flaoting around and everythign is in a dataframe.
```

